---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r, setoptions, echo=FALSE,warning=FALSE}
library(knitr)
opts_chunk$set(options(scipen = 1, digits = 2), message = FALSE, warning=FALSE)
```


```{r}
library(Hmisc)
library(lattice)
```
## Loading and preprocessing 

####1. Loading data
```{r}
if(!file.exists("activity.csv")){
  unzip("activity.zip")
}
activity_data<-read.csv("activity.csv")

```

## What is mean total number of steps taken per day? 
#####1. Total number of steps per day ignoring missing values 
```{r}
stepsPerDay <- aggregate(steps~date, activity_data, sum)

```

#####2. Make a histogram of the total number of steps taken each day
```{r}

hist(stepsPerDay$steps, main=paste("Total steps taken per day"), xlab="Number of steps", col = "red")
```

#####3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean_stepsPerDay<-mean(stepsPerDay$steps)
median_stepsPerDay<-median(stepsPerDay$steps)

```
* The mean of the dataset is **`r mean_stepsPerDay`** and the median is **`r median_stepsPerDay`**. 

## What is the average daily activity pattern?
#####1. Calculate average steps per interval 
```{r}
average_stepsPerInterval <- aggregate(steps ~ interval, activity_data, mean)
```

#####2. Plot average number of steps per interval in each day 
```{r}
plot(average_stepsPerInterval$interval,average_stepsPerInterval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average number of Steps per interval in each day")
```

#####3. Which interval contains the maximum number of steps? 
```{r}
maximum_interval <- average_stepsPerInterval[which.max(average_stepsPerInterval$steps),1]
```
* The interval containing the maximum number of steps is **`r maximum_interval`**.

## Imputing missing values
#####1. Calculate and report total number of missing values in the dataset
```{r}
total_NAs <- sum(is.na(activity_data$steps))
```
* The total number of missing values in the dataset is **`r total_NAs`**.

#####2. create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
imputedActivityData<-activity_data
imputedActivityData$steps<-impute(activity_data$steps, fun = mean) # Missing data replaced by mean
```

#####3. Make a histogram of the total number of steps taken each day
```{r}
stepsPerDayImputed <- aggregate(steps~date, imputedActivityData, sum)
hist(stepsPerDayImputed$steps, main=paste("Total steps taken per day"), xlab="Number of steps", col = "red")
```

#####4. Calculate and report the mean and median total number of steps taken per day
```{r}
mean_stepsPerDayImputed<-mean(stepsPerDayImputed$steps)
median_stepsPerDayImputed<-median(stepsPerDayImputed$steps)

#Calculate the difference of mean and median of the imputed and non-imputed datasets
meanDifference<-mean_stepsPerDayImputed - mean_stepsPerDay
medianDifference<-median_stepsPerDayImputed - median_stepsPerDay

#Impact of imputing missing data on total daily number of steps
stepsDifference<-sum(stepsPerDayImputed$steps) - sum(stepsPerDay$steps)
```
* The mean of the imputed dataset is **`r mean_stepsPerDayImputed`** and the median is **`r median_stepsPerDayImputed`**.
* The difference between the imputed data mean and non-imputed data mean is **`r meanDifference`**.Thus, the mean remains unchanged. 
* The difference between the imputed data median and non-imputed data median is **`r medianDifference`**.
* The difference between the total number of steps in the imputed and non-imputed data is `r stepsDifference`.Therefore, the imputed data has **`r stepsDifference`** more steps. 

## Are there differences in activity patterns between weekdays and weekends?
#####1. Create a new factor variable in the dataset with two levels:"weekday" and "weekend"
```{r}
imputedActivityData$dayType<-ifelse(as.POSIXlt(imputedActivityData$date)$wday %in% c(0,6), "Weekend", "weekday")
```
#####2. Make a panel plot of the 5-minute interval and average number of steps taken
```{r}
average_imputedActivityData<-aggregate(steps~interval+dayType, imputedActivityData, mean)
xyplot(average_imputedActivityData$steps~average_imputedActivityData$interval|average_imputedActivityData$dayType, main="Average Steps per Day by Interval",xlab="5-miute Interval", ylab="Average Number of Steps",layout=c(1,2), type="l")
```






